---
title: "Data Review"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = '')
library(tidyverse)
```

<style>body{text-align: justify;}</style>

## Data Import

```{r data_import,cache = TRUE}
if (!file.exists("leaf.zip")) {
  download.file("https://archive.ics.uci.edu/ml/machine-learning-databases/00288/leaf.zip","leaf.zip")
}
unzip("leaf.zip")
leaf <- read_csv("leaf.csv",col_types = "fddddddddddddddd")
leaf <- tibble(leaf)
names(leaf) <- c("Class","Specimen Number","Eccentricity","Aspect Ratio","Elongation","Solidity",
                 "Stohastic Convexivity","Isoperimetric Factor","Maximal Identation Depth","Lobedness",
                 "Average Intensity","Average Contrast","Smoothness","Third Moment","Uniformity","Entropy")
```

The data is already cleaned with no missing values and the codebook for references and features' explanations can 
be found in the <text style = "color:red">ReadMe.pdf</text> file. For the classification metrics of the classes the 
images in the __BW__ (black & white) folder were used.

## Part 1

We are going to group data by class and create a simple barplot to see how well represented with respect to samples 
the classes are.

```{r data_1}
fig_dat1 <- leaf %>% group_by(Class) %>% count
fig_dat1
```

## Part 2

In this section we will create a boxplot with eccentricity in the y-axis and classes in the x-axis, since we were instructed 
by experts who trained a linear model which used as predictors all the features 
that _Eccentricity_ is the most important factor.

```{r data_2}
fig_dat2 <- leaf %>% select(Class,Eccentricity) 
fig_dat2
```


## Part 3

The linear model which was used as a machine learning model for predictions, provided the information that Solidity and Entropy 
were the second and third most important factors, we are going to plot a scatter plot, coloring the points with respect to the 
class of the leaves.

```{r data_3}
fig_dat3 <- leaf %>% select(Class,Solidity,Entropy) 
fig_dat3
```

