---
title: "More ggplot Figures Peer Review"
output: html_document
---

<style>body{text-align: justify;}</style>

## Problem 1

Create the figure in the solution for Problem 1, using the data included in the R Markdown file.

```{r problem_1_data,include=FALSE}
#####DO NOT MODIFY THIS CHUNK
library(tidyverse)
set.seed(12345)
var1<-runif(50,0,10)
set.seed(12345)
var2<-var1+rnorm(50,5,2)
set.seed(12345)
var3<- var1*(-rnorm(50,1,.2))
dat1<-tibble(var1,var2,var3)
```

```{r problem_1,echo=FALSE,message=FALSE}
ggplot(dat1) + geom_point(aes(var1,var2),cex = 2) + geom_smooth(aes(var1,var2)) + 
  labs(x = "Variable 1",y = "Variable 2")
```

## Problem 2

Create the figure in the solution for Problem 2, using the data included in the R Markdown file.

```{r problem_2_data,include=FALSE}
#####DO NOT MODIFY THIS CHUNK
set.seed(10)
var1<-runif(50,0,10)
set.seed(10)
var2<-var1+rnorm(50,5,2)
set.seed(10)
var3<- var1*(-rnorm(50,1,.2))
dat2<-tibble(var1,var2,var3)
```


```{r solution_p2,echo=FALSE,message=FALSE}
library(GGally)
point_fun <- function(data,mapping,...) {
  ggplot(data = data) + geom_point(mapping = mapping,colour = "dark orange",cex = 2)
}
density_fun <- function(data,mapping,...) {
  ggplot(data = data) + geom_density(mapping = mapping,fill = "blue",color = "black")
}
ggpairs(data = dat2,lower = list(continuous = point_fun),diag = list(continuous = density_fun),
        upper = list(continuous = wrap("cor",family = "mono",size = 4)))
```

## Problem 3

Create the figure in the solution for Problem 3, using the data included in the R Markdown file.

```{r problem_3_data,include=FALSE}
#####DO NOT MODIFY THIS CHUNK
set.seed(12)
var1<-runif(50,0,10)
set.seed(12)
var2<-var1+rnorm(50,5,2)
set.seed(12)
var3<- var1*(-rnorm(50,1,.2))
dat3<-tibble(var1,var2,var3)
```

```{r problem3,echo=FALSE}
library(ggcorrplot)
r <- cor(dat3)
ggcorrplot(corr = r,type = "lower",title = "Correlation") + 
  theme(title = element_text(colour = 'grey',size = 16))
```

## Problem 4

Create the figure in the solution for Problem 4, using the data included in the R Markdown file.

```{r problem_4_data,include=FALSE}
set.seed(5678)
var1<-rnorm(10,25,5)
names<-letters[1:10]
dat4<-tibble(names,var1)
```

```{r problem_4,echo=FALSE}
ggplot(data = dat4,aes(x = var1,y = reorder(names,var1))) + geom_point(cex = 5) + 
  theme(panel.grid.major.y = element_line(linetype = "dotted",color = "red"),
        panel.grid.minor.x = element_blank(),panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank()) + 
  labs(x = "Variable 1",y = "")
```

## Problem 5

Create the figure in the solution for Problem 5, using the data included in the R Markdown file.

```{r problem_5_data,include=FALSE}
#####DO NOT MODIFY THIS CHUNK
set.seed(13)
var1<-rnorm(10,25,5)
names<-letters[1:10]
dat5<-tibble(names,var1)
```

```{r problem_5,echo=FALSE}
ggplot(data = dat5,aes(x = reorder(names,var1),y= var1)) + geom_segment(aes(xend = names, yend = 0),
                                                                        color = "dark green") +
  geom_point(color = "dark green") + labs(x = "",y = "Variable 1") + 
  theme(axis.text.x = element_text(angle = 90))
```

## Problem 6

Create the figure in the solution for Problem 6, using the data included in the R Markdown file.

```{r problem_6_data,include=FALSE}
#####DO NOT MODIFY THIS CHUNK
set.seed(8)
fiction<-tibble(Genre=rep("Fiction",5),Time=seq(1:5),Total=sample(1:10,5))
set.seed(7)
biography<-tibble(Genre=rep("Biography",5),Time=seq(1:5),Total=sample(1:10,5))
set.seed(9)
mystery<-tibble(Genre=rep("Mystery",5),Time=seq(1:5),Total=sample(1:10,5))
books_checked_out<-bind_rows(fiction,biography,mystery)
```

```{r problem_6,echo=FALSE}
ggplot(data = books_checked_out) + geom_area(position = "stack",aes(x = Time,y = Total,fill = Genre)) + 
  ggtitle("Books Checked Out")
```

## Problem 7

Create the figure in the solution for Problem 7, using the data included in the R Markdown file.

```{r problem_7_data,include=FALSE}
#####DO NOT MODIFY THIS CHUNK
books_checked_out2 <- books_checked_out %>%
  filter(Time==1 | Time==5) %>%
  pivot_wider(names_from = Time,values_from=Total) %>%
  rename(Time1=`1`,Time5=`5`)
```
   
```{r,problem_7,echo=FALSE,message=FALSE}   
library(ggalt)
ggplot(books_checked_out2) + 
  geom_dumbbell(aes(y = reorder(Genre,Time1),x = Time1,xend = Time5),colour_x = "red",
                                           colour_xend = "purple",size_x = 3.5,size_xend = 3.5) +
  ylab("Genre") + xlab("")
```

## Problem 8

Create the figure in the solution for Problem 8, using the data included in the R Markdown file.

```{r, problem_8_data,include=FALSE}
#####DO NOT MODIFY THIS CHUNK
pie_dat<-c(1,2,3,4,5)
```

```{r, problem_8,echo=FALSE}
###HINT: check the Help file for pie to see how to specify the color to the my_color palette
library(RColorBrewer)
my_color<-brewer.pal(5,"Spectral")
pie(pie_dat,col = my_color)
```


