---
title: "Week 3 Peer Review"
output: html_document
---

```{r setup, include=FALSE}

######DO NOT MODIFY. This will load required packages and data.

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

cces <- drop_na(read_csv("../Data/cces.csv"))

cel <- drop_na(read_csv("../Data/cel_volden_wiseman.csv"))

```

<style>body{text-align: justify;}</style>

Following the directions on the Coursera assignment page, you will make four original visualizations.
Note that the data for the CCES and CEL data are imported in code in the R Markdown file.

### Put your name here: GP

## Exercise 1

Explain what you are visualizing here:

We created a data frame with two columns each containing numbers from a normal distribution and then
plotted a scatter plot x-y, a text annotation and a linear fit to show that a regression line always
passes from the mean of $\bar{x}$ and $\bar{y}$ of the distributions.

Put your figure here:

```{r exercise1,echo=FALSE}
ggplot(data = data.frame(x = rnorm(500,0,4),y = rnorm(500,0,4)),aes(x,y)) + 
  geom_point(colour = "red",alpha = 0.3) + geom_smooth(method = 'lm',formula = "y ~ x") + 
  geom_point(data = data.frame(x=0,y=0),aes(x,y),cex = 5,colour = "black") + theme_bw() + 
  ggtitle("Regression of N(0,4) ~ N(0,4)")
```

## Exercise 2

Explain what you are visualizing here:

We want to plot the distribution of the numbers of how far away is the member from the majority median
DW-Nominate score indicative member ideology, with higher indicating a more conservative view, the data
used is from the center for effective lawmaking.

Put your figure here:

```{r exercise2,echo=FALSE}
ggplot(data = cel,aes(meddist)) + 
  geom_density(colour = "black",lwd = 2) +  
  geom_histogram(aes(y = ..density..),bins = 30,colour = "black",fill = "purple") + theme_bw() + 
  xlab("distance from median DW-Nominate Score") + ggtitle("Histogram With Density Plot")
```

## Exercise 3

Explain what you are visualizing here:

We would like to picture a barplot of votes per region based on a research of the public of opinion of
whether the interest rates of mortgages should be reduced. The data we will use is from the cooperative
congressional election study.

Put your figure here:

```{r exercise3,echo=FALSE}
cces$CC18_325b <- recode(cces$CC18_325b,`1` = "Support",`2` = "Oppose")
cces$region    <- recode(cces$region,`1` = "Northwest",
	`2` = "Midwest",`3` = "South",`4` = "West")
ggplot(cces,aes(CC18_325b)) + geom_bar(aes(fill = CC18_325b)) + facet_grid(. ~ region) + 
  theme(legend.title = element_blank()) + xlab('') + theme_classic() + 
  ggtitle("Reduce the mortgage interest deduction? Opinions per region.")
```

## Exercise 4

Explain what you are visualizing here:

We would like to make a simple line plot to summarize the total number of bills per congress that
became laws, the data we used is from the center for effective lawmaking.

Put your figure here:

```{r exercise4,echo=FALSE}
ggplot(data = cel %>% group_by(congress) %>% 
         summarize(all_law = sum(all_law)),aes(x = congress,y = all_law)) +
  geom_line(lwd = 2) + geom_point(colour = "red",alpha = 0.3,cex = 4) + 
  xlab("number of congress") + ylab("cumulative number bills that became laws") + 
  ggtitle("Bills Per Congress That Became Laws")
```
