---
title: "Week 2 Peer Review"
output: html_document
---

```{r setup, include=FALSE}

######DO NOT MODIFY. This will load required packages and data.

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
cces <- drop_na(read_csv("../Data/cces.csv"))
```
<style>body{text-align: justify;}</style>

Your objective is to replicate these figures, created using the Cooperative Congressional Election
Study data. These figures are similar to those we completed in the lecture videos.

### Put your name here: GP

## Exercise 1

Hints:

* For the x-axis, use the variable "ideo5".

* Make sure you recode the data for the "ideo5" variable to generate the correct names for the x-axis.
You will want to consult the codebook.

* Use the fill aesthetic to have R fill in the bars. You do not need to set the colors manually.

* Use guides() to drop the legend.

* Make sure the axis labels and figure title are correct.


```{r exercise1,echo=FALSE}
cces$ideo5 <- recode(cces$ideo5,`1` = "Very liberal",
                     `2` = "Liberal",`3` = "Moderate",`4` = "Conservative",
                     `5` = "Very conservative")
ggplot(data = cces) + geom_bar(aes(x = ideo5,fill = ideo5)) + guides(fill = FALSE) + 
  xlab("Ideology")
```

## Exercise 2

Hints:

* For the x-axis, use the variable "pew_religimp".

* Make sure you recode the data for the "pew_religimp" variable to generate the correct labels for the
x-axis. You will want to consult the codebook.

* Rename the column for Ideology to make sure the first letter is upper-case 
(to make the legend appear correctly).

* Use the fill aesthetic to have R fill in the bars. You do not need to set the colors manually.

* Make sure the axis labels and figure title are correct.

```{r exercise2,echo=FALSE}
cces$pew_religimp <- recode(cces$pew_religimp,`1` = "Very important",
	`2` = "Somewhat important",`3` = "Not too important",`4` = "Not at all important")
cces <- rename(cces,"Ideology" = "ideo5")
ggplot(cces,aes(x = pew_religimp)) + geom_bar(position = "stack",aes(fill = Ideology)) + 
  xlab("Importance Of Religion") + ggtitle("Importance Of Religion By Ideology")
```

## Exercise 3

Instructions:

* For this visualization, you are creating your own data for practice.

* Create a tibble/data frame with three columns: Semester, Student, and Grade.

* There should be six semesters and three students (Amanda, Betty, and Carol)

* Create grades for the students using the runif() command, with values between 80 and 100. 
Hint: you'll need 18 grades total.

The figure should look approximately like this (your vaules will be slightly different):


```{r exercise3,echo=FALSE}
set.seed(100)
grades_data <- data.frame(
  Semester = rep(1:6,each = 3),
  Students = rep(c("Amanda","Betty","Carol"),6),
  Grade   = runif(18,80,100)
)
ggplot(grades_data,aes(x = Semester,y = Grade)) + geom_line(aes(colour = Students)) + 
  facet_wrap(~Students) + theme(legend.position = "")
```
